pirouette: the error BEAST2 makes in inferring a phylogeny

### 2019-02-28 article.tex
Create all ESSes and evidences and figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | r | l | l | l | l | }
\hline
\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
\cline{2-5}
                                  & \textbf{HPD} & \textbf{range} & \textbf{HPD} & \textbf{range} \\
\hline
1  & $0.20$  & $0.0 - 0.27$  &         &               \\
2  & $0.02$  & $0.0 - 0.08$  &         &               \\
3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
\hline
\end{tabular}
\caption{Results of the examples. For the true and twin tree,
  the highest posterior density (HPC) and range of the errors 
  is shown. Experiment 4 is split in two rows
  by either the candidate ('C') or generative model ('G') 
}
\label{tab:results}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Here we discuss how to interpret the results of \verb;pirouette;.
Figure \ref{fig:example_1} shows that the generative model 
gives an error distribution with a highest posterior density (HPD) 
\giovanni{probably "median" is a better word}
\richel{HPC is a valid Bayesian term I learned from the BEAST book}
at 0.015
The range of errors goes from 0.0 to 0.097 
As we know the generative model, 
this error distribution is the minimum noise that BEAST2 produces 
for that generative model.
We can measure the effective samples as well (not shown), 
and all ESSes are above
the recommended value of 200 \cite{drummond2015bayesian}, 
as shown in table \ref{tab:esses_example_1}.

Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
with an HPD at 0.02. The range of errors goes from 0.0 to 0.08. 
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of $0.675$. 
This is (unexpectedly) better than the generative
model with a weight of $0.166$. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
This means that it is rational to use this inference model to 
infer a posterior phylogeny distribution, 
as is done by \verb;pirouette;.
Also here, the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_2}.

Figure \ref{fig:example_3} shows that the hand-picked model 
gives an error distribution with an HPD at 0.018, 
with a range from 0.0 to 0.1.
The twin tree has an error distribution with a highest density at 0.019,
with a range from 0.0 to 0.11, which is an indication of the
error made by BEAST on that tree prior.
As there is little difference between the true and twin tree,
the (unknown) diversification model has little impact on the error
BEAST2 makes, hinting that the constant-rate birth-death model is
sufficiently good enough to explain the true phylogeny. 
The diversification model apparently has minor impact and need not to be
added to BEAST2.

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | r | l | l | l | l | }
\hline
\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
\cline{2-5}
                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
\hline
1  & $0.02$  & $0.0 - 0.27$  &         &               \\
2  & $0.02$  & $0.0 - 0.08$  &         &               \\
3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
\hline
\end{tabular}
\caption{Results of the examples. 
  For the true and twin tree,
  the median and range of the errors 
  is shown. Experiment 4 is split in two rows
  by either the candidate ('C') or generative model ('G') 
}
\label{tab:results}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{check numbers}

Here we discuss how to interpret the results of \verb;pirouette;.
Figure \ref{fig:example_1} shows that the generative model 
gives an error distribution with a median of \richel{?}
The range of errors goes from \richel{?} to \richel{?}.
As we know the generative model, 
this error distribution is the minimum noise that BEAST2 produces 
for that generative model.
We can measure the effective samples as well (not shown), 
and all ESSes are above
the recommended value of 200 \cite{drummond2015bayesian}, 
as shown in table \ref{tab:esses_example_1}.

Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
with a median of \richel{?}. 
The range of errors goes from \richel{?} to \richel{?}. 
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of $0.675$. 
This is (unexpectedly) better than the generative
model with a weight of $0.166$. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
This means that it is rational to use this inference model to 
infer a posterior phylogeny distribution, 
as is done by \verb;pirouette;.
Also here, the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_2}.

Figure \ref{fig:example_3} shows that the hand-picked model 
gives an error distribution with a median of \richel{?}, 
The twin tree has an error distribution with a median of \richel{?},
with a range from 0.0 to \richel{?}, which is an indication of the
error made by BEAST on that tree prior.
As there is little difference between the true and twin tree,
the (unknown) diversification model has little impact on the error
BEAST2 makes, hinting that the constant-rate birth-death model is
sufficiently good enough to explain the true phylogeny. 
The diversification model apparently has minor impact and need not to be
added to BEAST2.
We can draw this conclusion with confidence, as the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_3}.

Figure \ref{fig:example_4} shows that \richel{?}
and 
gives an error distribution with a median of \richel{?}, 
The twin tree has an error distribution with a median of \richel{?},
with a range from 0.0 to \richel{?}, which is an indication of the
error made by BEAST on that tree prior.
The inference model with the highest evidence follows a JC site model, 
RLN clock model and BD tree prior, with a weight of ?. 
The complete list of all evidences is shown in table \ref{tab:evidences_example_4}.
As there is little difference between the true and twin tree,
the (unknown) diversification model has little impact on the error
BEAST2 makes, hinting that the constant-rate birth-death model is
sufficiently good enough to explain the true phylogeny. 
The diversification model apparently has minor impact and need not to be
added to BEAST2.
We can draw this conclusion with confidence, as the ESSes are well above 200, 
as shown in table \ref{tab:esses_example_4}.

@@

\richel{
  I'd enjoy to put all intermediate output here: alignments, twin trees,
  posteriors, nLTTs, errors. Now we only show the ones relevant in the text
}

@@

\input{example_3_esses.latex}
% has label tab:esses_example_3

\input{example_4_esses.latex}
% has label tab:esses_example_4

\input{example_4_evidences.latex}
% has label tab:evidences_example_4
